An Embedding Framework for Consistent Polyhedral Surrogates

### 2021-06-21 poly-embed-journal/poly-embed-journal.tex
Pre meeting touchups.  Partially through Claim 9 but not done

\subsection{Phase 1: construct $\V^*$}

@@

\begin{itemize}
\item Define $G = \{(x,c) \mid c \leq g(x)\} \subseteq \reals^\Y \times \reals$ to be the hypograph of $g$.
\item We define $H^+_v := H^+_{([v;-1], 0)}$ for a $v \in \reals^\Y_+$, where $[v; -1]\in\reals^\Y_+\times\reals$.
\item Additionally, denote $H_y^+ := H_{([e_y; 0], 0)}^+$. % \{ (x,c) \in \reals^\Y \times \reals \mid x_y \geq 0\}$.
\item Again, we let $H_v := H_{([v;-1], 0)}$ for $v \in \reals^\Y_+$.

\item Define $F_v := H_v \cap G$ for any $v\in\reals^\Y_+$.

@@

\begin{claim}
	$G = \cap \H$.
\end{claim}
\begin{proof}
	$(x,c) \in G \iff g(x) - c \geq 0 \iff \min_{v \in \V}\inprod{v}{x} - c \geq 0$ and $x \in \reals_+^\Y$, which is true if and only if $\inprod{v}{x} - c \geq 0 \,\forall v \in \V$ and $x_y \geq 0$ for all $y$.
	In turn, this statement holds if and only if $(x,c) \in H^+_v$ for all $v \in \V$ and in $H^+_y$ for all $y \in \Y$, so $(x,c) \in \cap \H$.
\end{proof}

@@

\begin{claim}
	Claim: $G$ is $(|\Y|+1)$-dimensional.  %\raf{$g$ is nonnegative on $\reals^\Y_+$, so $G$ contains $\{(x,c) \mid x\in\reals^\Y_+, c \leq 0\}$, which is full-dim.}
\end{claim}
\begin{proof}
	Since $g$ is nonnegative on $\reals_+^\Y$, $G$ therefore contains $\{(x,c) \mid x\in\reals^\Y_+, c \leq 0\}$, which is $(|\Y| + 1)$-dimensional.
\end{proof}

@@

\begin{claim}
	$\H_\Y \subseteq \H^*$.
\end{claim}
\begin{proof}
	We argue $\H_\Y \subseteq \H^*$; if there was a $y \in \Y$ such that $H_y = \{(x,c) \mid x_y \geq 0\}$ was not in $\H^*$, then we would either have some $c_1 > 0$ such that $\{(x,c) \mid x_y \geq c_1\} \in \H^*$, or we have a point $x$ such that $x_y < 0$ but $g(x) > -\infty$.
	The first cannot happen as we take $f$ to be defined (and finite) on $p = \delta_y \in \simplex$, and $f$ and $g$ match on the domain of $f$.
	Moreover, the second cannot be true by construction of $g$.	
\end{proof}

@@

@@

\begin{claim}
	For all $x\in\reals^\Y_+$, there exists $v\in\V^*$ such that $H_v$ supports $G$ at $(x,g(x))$.
\end{claim}
\begin{proof}
We can write any face $F_v = H_v \cap G = \{x \in \reals_+^{d-1} \mid \inprod{x}{v} = g(x)\}$.
Since $H_v^+$ must be a valid inequality to define a face, we know that $G \subseteq H_v^+$.
If there was no $v \in \V^*$ such that $x \in F_v$, 
Since $\H_\V = \H_\Y \cup \H_{\V^*}$, if there was no $v \in \V^*$ such that $x \in F_v$, it must be the case that $\min_{H_v \in \H_\Y} \inprod{v}{x} = \min_{y \in \Y} x_y = 0 = g(x)$, and we must have $x \not \in \reals_{++}^\Y$.

\jessie{If $G$ is not locally constant}
Since $g$ is continuous (as it is convex), and not constant \jessie{??}, there must be some $x' \in B(x,\epsilon) \cap \reals_+^\Y$ such that $g(x') > 0$ and therefore no $y \in \Y$ would define a face $F_{v^*}$ supporting $G$ at $x'$.
That $v^*$ must then be in $\V^*$, and we claim $\inprod{v^*}{x} = 0$, which would imply $F_{v^*}$ also supports $G$ at $x$.
If $\inprod{v^*}{x} > 0 = g(x)$ (recall $g$ nonnegative on $\reals_+^\Y$) then we contradict continuity of $g$ on $\reals_+^\Y$. \jessie{Not the level of rigor that we want.  Also using lots of polyhedral structure here.}
%First, $\inprod{v^*}{x} \geq 0$ by definition of $g$ nonnegative on $\reals_+^\Y$, and $\leq 0$

\jessie{Claim: $x$ must be on boundary of $\reals_+^\Y$ and therefore $g(x) = 0$, and we have a matchy matchy with a $v^* \in \V^*$.}

\jessie{If $x \in \reals_{++}^\Y$, then}

@@

  Observe that $x \in \reals^Y_+ \iff (x, c) \in \cap \H_\Y$.
  Moreover,
  \begin{align*}
  (x,c) \in G &\iff g(x) \geq c \\
  &\iff \inprod{v}{x} \geq c \,\, \forall v \in \V \\
  &\iff (x,c) \in H^+_{v} \,\, \forall v \in L(\R) \\
  &\iff \inprod{L(r)}{x} \geq c \,\, \forall r \in \R\\
  &\iff (x,c) \in H^+_{L(r)} \,\, \forall r \in \R \\
  &\iff (x,c) \in \cap \H_\V \\ 
  \end{align*}
  Combining the two equalities, we have $G = \cap \H$.

@@

\begin{claim} 
For all $v\in\reals^\Y_+$, if $H_v$ supports $G$,  $\dim(\pi(F_v)) = \dim(F_v)-1$.
\end{claim}
\begin{proof}
	\jessie{TODO}
	Facts we can potentially use:
	$F_v = \H_v \cap G \neq \emptyset$ by assumption.
	$\dim(F_v) = \dim(\affhull(F_v))$
	$\dim(\pi(F_v)) = \dim(F_v) - 1$.
\end{proof}

@@

\begin{corollary}
	For $v,v^*$ such that $F_v \subseteq F_{v^*}$ and $v^*\in \V^*$, $\pi(F_{v}) \subseteq \pi(F_{v^*})$. \raf{This will show (3) once we translate back}
\end{corollary}
\begin{claim}
	$\R'$ is a representative set for $L$ if and only if $\V \subseteq L(\R')$.
	\jessie{Is this true?  You could have hinge with a really bad loss vector that is never uniquely optimal (ex: $u = 2$) and $(3,0) \in L(\R)$ but not in $L(\{-1,1\})$.}
\end{claim}

@@

Let $f:\reals^\Y\to\reals_+\cup\{-\infty\}$ be a polyhedral concave function with $\dom(f) = \simplex$.
\begin{claim}
	We may write $f(p) = \min_{v\in\V} \inprod{p}{v}$ for some finite set $\V \subseteq \reals^\Y_+$ \raf{show you can do this using the $h$ and $\delta$ representation from Rockafellar -- argue that because $p\in\simplex$ we can write $\inprod{p}{b_i}-\beta_i = \inprod{p}{b_i-\beta_i\ones}$.}
\end{claim}
\begin{proof}
  We will think of $f$ as defined $f:\reals^\Y\to\reals_+\cup\{-\infty\}$  with $\dom(f) = \simplex$.
  \raf{Make formal}
  For $p \in \simplex$, we know $\sum_i p_i = 1$, and can write any inner product $\inprod{p}{b} - \beta = \inprod{p}{b} - \inprod{p}{\beta \ones} = \inprod{p}{b - \beta \ones}$.
  If $p \not \in \simplex$, then $f(p) = -\infty$ and inner products are not used to compute $F$.
  Moreover, since $f$ is polyhedral, it is finitely generated \citep[Proposition 19.1.2]{rockafellar1997convex} and can be written 
	\begin{align*}
	f(p) &= h(p) - \delta(p \mid C)\\
		 &= \min(\inprod{p}{b_1} - \beta_1, \ldots, \inprod{p}{b_k} - \beta_k) - \delta(p \mid \simplex)\\
		 &= \min(\inprod{p}{b_1 - \beta_1 \ones}, \ldots, \inprod{p}{b_k - \beta_k \ones}) - \delta(p \mid \simplex)~.~
	\end{align*}
\end{proof}

Note: $\theta(v) = \pi(F_v) \cap \simplex$. \raf{But haven't def $\theta$ yet}

Argue $\theta(v)$ is full-dim in $\simplex$ if and only if $\pi(F_v)$ is $|\Y|$-dim, if and only if $F_v$ is a facet. \raf{(2)}
\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

>>>>

\subsection{Phase 1: construct $\V^*$}
Suppose we have a finite set $\V \subset \reals_+^d$.

@@

\begin{itemize}
\item Define $G = \{(x,c) \mid c \leq g(x)\} \subseteq \reals^d \times \reals$ to be the hypograph of $g$.
\item We define $H^+_v := H^+_{([v;-1], 0)}$ for a $v \in \reals^d_+$, where $[v; -1]\in\reals^d_+\times\reals$.
\item Additionally, denote $H_y^+ := H_{([e_y; 0], 0)}^+$ for any $y \in [d]$. % \{ (x,c) \in \reals^\Y \times \reals \mid x_y \geq 0\}$.
\item Again, we let $H_v := H_{([v;-1], 0)}$ for $v \in \reals^d_+$.

\item Define $F_v := H_v \cap G$ for any $v\in\reals^d_+$.

@@

\begin{claim}
	$G = \cap \H$.
\end{claim}
\begin{proof}
	$(x,c) \in G \iff g(x) - c \geq 0 \iff \min_{v \in \V}\inprod{v}{x} - c \geq 0$ and $x \in \reals_+^d$, which is true if and only if $\inprod{v}{x} - c \geq 0 \,\forall v \in \V$ and $x_y \geq 0$ for all $y$.
	In turn, this statement holds if and only if $(x,c) \in H^+_v$ for all $v \in \V$ and in $H^+_y$ for all $y \in \Y$, so $(x,c) \in \cap \H$.
\end{proof}

@@

\begin{claim}
	Claim: $G$ is $(d+1)$-dimensional.  %\raf{$g$ is nonnegative on $\reals^\Y_+$, so $G$ contains $\{(x,c) \mid x\in\reals^\Y_+, c \leq 0\}$, which is full-dim.}
\end{claim}
\begin{proof}
	Since $g$ is nonnegative on $\reals_+^\Y$, $G$ therefore contains $\{(x,c) \mid x\in\reals^\Y_+, c \leq 0\}$, which is $(|\Y| + 1)$-dimensional.
\end{proof}

@@

\begin{claim}
	$\H_\Y \subseteq \H^*$.
\end{claim}
\begin{proof}
	We argue $\H_\Y \subseteq \H^*$; if there was a $y \in \Y$ such that $H_y = \{(x,c) \mid x_y \geq 0\}$ was not in $\H^*$, then we would either have some $c_1 > 0$ such that $\{(x,c) \mid x_y \geq c_1\} \in \H^*$, or we have a point $x$ such that $x_y < 0$ but $g(x) > -\infty$.
	The first cannot happen as we take $g$ to be defined (and finite) on $x = e_y \in \reals^d_+$.
	Moreover, the second cannot be true by construction of $g$ including the $0-\infty$ indicator on $\reals^d_+$.	
\end{proof}

@@

Since $G$ is full-dimensional, all facets of $G$ are uniquely determined by some $H^+ \in \H^*$.
Any facet $F$ must then be some intersection of an $H_y \cap G$ or $H_v \cap G$.
Take $\V^*$ to be the set of such $H_v$ hyperplanes.
%Since $\H_\Y \subseteq \H^*$, we can consider $\H_{\V^*} := \H \setminus \H_{\Y}$, and $\V^*$ to be the set generating $\H_{\V^*}$.

@@

\begin{claim}
	For all $x\in\reals^\Y_+$, there exists $v^*\in\V^*$ such that $H_{v^*}$ supports $G$ at $(x,g(x))$.
\end{claim}
\begin{proof}
We can write any face $F_v = H_v \cap G = \{x \in \reals_+^{d-1} \mid \inprod{x}{v} = g(x)\}$.
Since $H_v^+$ must be a valid halfspace to define a face, we know that $G \subseteq H_v^+$.
%If there was no $v \in \V^*$ such that $x \in F_v$, 
Since $\H^* = \H_\Y \cup \H_{\V^*}$, if there was no $v \in \V^*$ such that $x \in F_v$, it must be the case that $\min_{H_v \in \H_\Y} \inprod{v}{x} = \min_{y \in \Y} x_y = g(x) = 0$ by construction of the halfspaces in $\H_\Y$. %, and we must have $x \not \in \reals_{++}^\Y$.

\jessie{If $g$ is not locally constant}
Since $g$ is continuous on $\reals^d_+$ (as it is convex), and not constant \jessie{??}, there must be some $x' \in B(x,\epsilon) \cap \reals_+^\Y$ such that $g(x') > 0$ and therefore no $y \in \Y$ would define a face $F_{v^*}$ supporting $G$ at $x'$.
(We know such a face exists as $G$ is polyhedral and finite.)
That $v^*$ must then be in $\V^*$, and we claim $\inprod{v^*}{x} = 0$, which would imply $H_{v^*}$ also supports $G$ at $x$.
If $\inprod{v^*}{x} > 0 = g(x)$ (recall $g$ nonnegative on $\reals_+^\Y$) then we contradict continuity of $g$ on $\reals_+^\Y$. 
Therefore, we must have $\inprod{v^*}{x} = 0$, and thus $H_{v^*}$ also supports $G$ at $x$.
\jessie{Not the level of rigor that we want.  Also using lots of polyhedral structure here.}
%First, $\inprod{v^*}{x} \geq 0$ by definition of $g$ nonnegative on $\reals_+^\Y$, and $\leq 0$

%\jessie{Claim: $x$ must be on boundary of $\reals_+^\Y$ and therefore $g(x) = 0$, and we have a matchy matchy with a $v^* \in \V^*$.}

%\jessie{If $x \in \reals_{++}^\Y$, then}

@@

  Observe that $x \in \reals^\Y_+ \iff (x, c) \in \cap \H_\Y$.
  Moreover,
  \begin{align*}
  (x,c) \in G &\iff g(x) \geq c \\
  &\iff \inprod{v}{x} \geq c \,\, \forall v \in \V \\
  &\iff (x,c) \in H^+_{v} \,\, \forall v \in L(\R) \\
  &\iff \inprod{L(r)}{x} \geq c \,\, \forall r \in \R\\
  &\iff (x,c) \in H^+_{L(r)} \,\, \forall r \in \R \\
  &\iff (x,c) \in \cap \H_\V \\ 
  \end{align*}
  Combining the two equalities (e.g., $\cap \H = (\cap \H_\Y) \cap (\cap \H_\V)$), we have $G = \cap \H$.

@@

\begin{claim} 
For all $v\in\reals^\Y_+$, if $H_v$ supports $G$,  $\dim(\pi(F_v)) = \dim(F_v)-1$.
\end{claim}
\begin{proof}
	By assumption, $G$ is nowhere locally constant, meaning $([\vec 0; -1], b) \not \in \V^*$ for any $b \in \reals$.
	Therefore, the dimension of $\pi(F_v) = \pi(H_v \cap G) = \pi(\{(x,c) \in G \mid \inprod{v}{x} = c, c = g(x) \}) = \{x \in \reals^d_+ \mid \inprod{v}{x} = c\}$.
	Any $k\leq d$ vectors $\{z_i\}_{i=1}^k$ forming a basis for $\{(x,c) \in G \mid \inprod{v}{x} = c, c = g(x) \}$ \jessie{....}

	\jessie{TODO}
	Facts we can potentially use:
	$F_v = \H_v \cap G \neq \emptyset$ by assumption.
	$\dim(F_v) = \dim(\affhull(F_v))$
	$\dim(\pi(F_v)) = \dim(F_v) - 1$.
\end{proof}

@@

\begin{corollary}
	For $v,v^*$ such that $F_v \subseteq F_{v^*}$ and $v^*\in \V^*$, $\pi(F_{v}) \subseteq \pi(F_{v^*})$. \raf{This will show (3) once we translate back}
\end{corollary}

\begin{claim}
	\jessie{This should probably move down to translating for properties since representative sets are at the core, about properties}
	$\R'$ is a representative set for $L$ if and only if $\V^* \subseteq L(\R')$.
\end{claim}
\begin{proof}
	Fix $v^* \in \V^*$.
	If $\R'$ is representative, then for any $p \in \simplex$ such that , 
	As $g$ is $1$-homogeneous and minimizable, for every $v^* \in \V^*$, there must be some $p \in \simplex$ such that $\inprod{v^*}{p} = \inf_v \inprod{v}{x}$.
	Thus, it suffices to show subset inclusion for any $p \in \simplex$.
	Moreover, since $\R'$ is representative, there is some $r' \in \R'$ such that $\inprod{L(r')}{p} = \inf_{v} \inprod{v}{x}$.

	For the reverse implication, $\V^*$ is representative (since $\cup_{v^* \in \V^*} F_{v^*} = \reals_+^\Y$ in Corollary 8\jessie{make reference}), so any set containing $\V^*$ is also representative.
\end{proof}

@@

Let $f:\reals^\Y\to\reals_+\cup\{-\infty\}$ be a polyhedral concave function with $\dom(f) = \simplex$.
\jessie{Want to construct so that it matches $g$}
\begin{claim}
	We may write $f(p) = \min_{v\in\V} \inprod{p}{v}$ for some finite set $\V \subseteq \reals^\Y_+$ \raf{show you can do this using the $h$ and $\delta$ representation from Rockafellar -- argue that because $p\in\simplex$ we can write $\inprod{p}{b_i}-\beta_i = \inprod{p}{b_i-\beta_i\ones}$.}
\end{claim}
\begin{proof}
  We will think of $f$ as defined $f:\reals^\Y\to\reals_+\cup\{-\infty\}$  with $\dom(f) = \simplex$.
  \raf{Make formal}
  For $p \in \simplex$, we know $\sum_i p_i = 1$, and can write any inner product $\inprod{p}{b} - \beta = \inprod{p}{b} - \inprod{p}{\beta \ones} = \inprod{p}{b - \beta \ones}$.
  If $p \not \in \simplex$, then $f(p) = -\infty$ and inner products are not used to compute $F$.
  Moreover, since $f$ is polyhedral, it is finitely generated \citep[Proposition 19.1.2]{rockafellar1997convex} and can be written 
	\begin{align*}
	f(p) &= h(p) - \delta(p \mid C)\\
		 &= \min(\inprod{p}{b_1} - \beta_1, \ldots, \inprod{p}{b_k} - \beta_k) - \delta(p \mid \simplex)\\
		 &= \min(\inprod{p}{b_1 - \beta_1 \ones}, \ldots, \inprod{p}{b_k - \beta_k \ones}) - \delta(p \mid \simplex)~.~
	\end{align*}
\end{proof}

\begin{claim}
  $\theta(v) = \pi(F_v) \cap \simplex$. \raf{But haven't def $\theta$ yet}
\end{claim}

Argue $\theta(v)$ is full-dim in $\simplex$ if and only if $\pi(F_v)$ is $|\Y|$-dim, if and only if $F_v$ is a facet. \raf{(2)}

\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
